### 2013-12-04 manuscripts/reply/reply-figures.tex
Working on reply ms.
Moving several figures to Supporting Information. Miscellaneous
edits throughout.

%% Power results
\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
    estimate the dispersion index of divergence times (\vmratio{}) from
    analyses of simulated datasets where \divt{} for 22 pairs of populations is
    drawn from a series of uniform distributions, $\divt{} \sim U(0,
    \divt{max})$.
    The proportion of estimates less than the true value of,
    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
    Each plot represents unadjusted median estimates from 500 simulation
    replicates analyzed using $5\e6$ samples from the prior.
}{figPowerAccOmegaMedian}

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
    estimate the dispersion index of divergence times (\vmratio{}) from
    analyses of simulated datasets where \divt{} for 22 pairs of populations is
    drawn from a series of uniform distributions, $\divt{} \sim U(0,
    \divt{max})$.
    The proportion of estimates less than the true value of,
    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
    Each plot represents GLM-regression-adjusted mode estimates from 500
    simulation replicates analyzed using $5\e6$ samples from the prior.
}{figPowerAccOmegaModeGLM}

@@

%% Validation results
\mFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
    An assessment of the approximate Bayesian model-averageing approach of
    \citet{Hickerson2013} under the ideal conditions when the prior model
    is correct (i.e., the pseudo-replicate datasets are simulated from
    parameters drawn from the same prior distributions used in the
    analysis).
    The plots show the relationship between the estimated posterior and true
    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
    on 50,000 simulations.
    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
    posterior estimate from each simulation replicate.
    The prior settings for all replicates included five prior models with
    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
    U(0.0001, 0.05)$ for all five models, and
    $M_1: \divt{} \sim U(0, 0.1)$,
    $M_2: \divt{} \sim U(0, 1)$,
    $M_3: \divt{} \sim U(0, 5)$,
    $M_4: \divt{} \sim U(0, 10)$, and
    $M_5: \divt{} \sim U(0, 20)$.
    The number of samples from the prior was $2.5\e6$.
    The simulated data structure was 8 population pairs, with a single 1000
    bp locus sampled from 10 individuals from each population.  The 50,000
    estimates of the posterior probability of one divergence event were
    assigned to 20 bins of width 0.05.
    The estimated posterior probability of each bin is plotted against the
    proportion of replicates in that bin with a true value consistent with
    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
}{figValidationMCBehavior}

@@

%:FIGURE-saturation plot
\mFigure{../../saturation/saturation-plot.pdf}{
    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
    \citep{Takahata1985} as a function of divergence time between populations.
    Each plot represents 1100 pairs of parameter draws and summary statistics
    calculated from the simulated data.
    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
    U(0.0005, 0.02)$.
}{figSaturationPlot}

>>>>

%% Power results

@@

@@

### 2013-12-04 manuscripts/reply/reply-si-figures.tex
Working on reply ms.
Moving several figures to Supporting Information. Miscellaneous
edits throughout.

>>>>

\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
    estimate the dispersion index of divergence times (\vmratio{}) from
    analyses of simulated datasets where \divt{} for 22 pairs of populations is
    drawn from a series of uniform distributions, $\divt{} \sim U(0,
    \divt{max})$.
    The proportion of estimates less than the true value of,
    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
    Each plot represents unadjusted median estimates from 500 simulation
    replicates analyzed using $5\e6$ samples from the prior.
}{figPowerAccOmegaMedian}

\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
    estimate the dispersion index of divergence times (\vmratio{}) from
    analyses of simulated datasets where \divt{} for 22 pairs of populations is
    drawn from a series of uniform distributions, $\divt{} \sim U(0,
    \divt{max})$.
    The proportion of estimates less than the true value of,
    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
    Each plot represents GLM-regression-adjusted mode estimates from 500
    simulation replicates analyzed using $5\e6$ samples from the prior.
}{figPowerAccOmegaModeGLM}

%% Validation results
\siFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
    An assessment of the approximate Bayesian model-averageing approach of
    \citet{Hickerson2013} under the ideal conditions when the prior model
    is correct (i.e., the pseudo-replicate datasets are simulated from
    parameters drawn from the same prior distributions used in the
    analysis).
    The plots show the relationship between the estimated posterior and true
    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
    on 50,000 simulations.
    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
    posterior estimate from each simulation replicate.
    The prior settings for all replicates included five prior models with
    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
    U(0.0001, 0.05)$ for all five models, and
    $M_1: \divt{} \sim U(0, 0.1)$,
    $M_2: \divt{} \sim U(0, 1)$,
    $M_3: \divt{} \sim U(0, 5)$,
    $M_4: \divt{} \sim U(0, 10)$, and
    $M_5: \divt{} \sim U(0, 20)$.
    The number of samples from the prior was $2.5\e6$.
    The simulated data structure was 8 population pairs, with a single 1000
    bp locus sampled from 10 individuals from each population.  The 50,000
    estimates of the posterior probability of one divergence event were
    assigned to 20 bins of width 0.05.
    The estimated posterior probability of each bin is plotted against the
    proportion of replicates in that bin with a true value consistent with
    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
}{figValidationMCBehavior}

%:FIGURE-saturation plot
\siFigure{../../saturation/saturation-plot.pdf}{
    The summary statistics $\pi$ \citep{Tajima1983} and $\pi_{net}$
    \citep{Takahata1985} as a function of divergence time between populations.
    Each plot represents 1100 pairs of parameter draws and summary statistics
    calculated from the simulated data.
    Prior settings for the simulations were $\divt{} \sim U(0, 20)$,
    $\meanDescendantTheta{} \sim U(0.0005, 0.04)$, and $\ancestralTheta{} \sim
    U(0.0005, 0.02)$.
}{figSaturationPlot}

### 2013-12-04 manuscripts/reply/short-reply.tex
Working on reply ms.
Moving several figures to Supporting Information. Miscellaneous
edits throughout.

%CWL: I would suggest removing this section
\section*{The power analysis of \citet{Hickerson2013}}
\citet{Hickerson2013} present a power analysis in which they find that
\msb has the power to infer multiple divergence events when divergence times
are random over hundreds of thousands of generations (rather than millions as
demonstrated by \citet{Oaks2012}).
It is not surprising that under certain prior conditions the method has
increased power to detect temporal variation in divergences over narrower time
windows than millions of generations.
However, what is important to consider is whether those conditions are relevant
to real-world applications of the method.
\citet{Oaks2012} explore the behavior of the method under three divergence time
priors, as narrow as 0--5 coalescent units.
This is quite narrow considering that this expresses the prior belief that all
22 taxon pairs diverged within this window.
\citet{Hickerson2013} limit their power analysis to a single prior of 0--1
coalescent unit.
This prior setting assumes that there is enough \emph{a priori} information to
be certain that all taxa (18 for their simulations) diverged within
the last 4$\globalpopsize$ generations.
This does not seem applicable to most empirical systems.
Furthermore, even when such prior information is available, the results of
\citet{Hickerson2013} demonstrate the method is limited to detecting multiple
divergences on the scale of hundreds of thousands of generations, assuming a
per-site rate of $1.92\e{-8}$ mutations per generation.
Assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), their
results demonstrate the method can reliably reject one divergence event when
divergences are random over a period of 3 million generations.

\section*{Model likelihoods or insufficient prior sampling as the cause of bias
in \msb}
% One of the points of disagreement between \citet{Hickerson2013} and
% \citet{Oaks2012} is the underlying mechanism causing the bias toward models
% with clustered divergences (i.e., models with few divergence events).
\citet{Oaks2012} present simulation-based analyses that suggest the broad
uniform prior on divergence-time parameters result in small marginal
likelihoods of models with more divergence-time parameters, and thus a
preference for models with few divergence parameters.
\citet{Hickerson2013} conclude that the bias is due to the inefficient
rejection algorithm implemented in \msb.

\citet{Hickerson2013} present an interesting probabilistic argument
 to show that insufficient prior sampling is to blame for the bias.
They argue the widest of the three priors used by \citet{Oaks2012} was very
unlikely to produce any prior samples with large numbers of divergence times
that were consistent with the Philippine data.
There are a few issues with this argument.
The probabilities they present assume the gene divergence times
estimated by \citet{Oaks2012} are correct.
The sole purpose of the estimates of ultrametric gene trees presented by
\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
times across the 22 taxa.
These analyses assumed an arbitrary strict mutation rate of $2\e{-8}$ for all
taxa, and are, of course, subject to estimation error.
Furthermore, the branch-length units of the gene trees are in millions of
years, whereas the divergence time prior of \msb is in generations, thus the
logic of \citet{Hickerson2013} requires the additional assumption that all 22
Philippine taxa have a generation time of one year.
Thus, the argument of \citet{Hickerson2013} that divergence time estimates of
\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
questionable, especially given our findings above on the behavior of \msb when
empirically informed uniform priors are employed.

Also, if we make all of these assumptions and assume the gene divergence times
are estimated without error, the probabilistic argument only applies to one of
the three different priors used by \citet{Oaks2012}.
The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5
mya)$) closely mirrors the range of estimates of gene-divergence times.
Applying Hickerson et al.'s (\citeyear{Hickerson2013}) probability equations
demonstrates this prior is densely populated with samples with large numbers of
divergence parameters that are consistent with the gene divergence estimates.
Thus, according to the argument of \citet{Hickerson2013}, if insufficient prior
sampling is to blame for the bias, it should be much reduced under the narrow
prior on \divt{}.
However, the magnitude of the bias is very similar across all three priors
explored by \citet{Oaks2012}.
\citet{Hickerson2013} point out a case where the narrow prior performs
slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
However, it is important to note that these results suffered from a bug
in \msb, and there are many cases after \citet{Oaks2012} corrected the 
bug where the narrow prior performs slightly worse (see panels D--J of
Figures 3 and S12).

To disentangle whether model likelihoods or insufficient prior sampling is to
blame for the biases revealed by \citet{Oaks2012}, we must look at the
different predictions made by these two phenomena.
One example, as discussed by \citet{Oaks2012}, is that insufficient prior
sampling should create higher variance in posterior estimates, and thus it
should cause analyses to be sensitive to the number of samples drawn from the
prior.
\citet{Oaks2012} do not see such sensitivity when they compare prior sample
sizes of $2\e6$, $5\e6$, and $10^7$.

To explore this prediction further, we repeat the analysis of the Philippines
dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
When we look at the trace of the estimates of the dispersion index of
divergence times (\vmratio{}) as the prior samples accumulate
(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
GLM-regression-adjusted estimates.
This suggests that insufficient prior sampling did not play a large
role in the bias found by \citet{Oaks2012}.

Our results presented above that demonstrate the bias of the model-averaging
approach of \citet{Hickerson2013} both toward models with narrower \divt{}
priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
\ref{figPowerOmegaProb}) strongly suggest the primary cause is the uniform
priors reducing the marginal likelihoods of large models.
In all of our model-averaging analyses, all prior models have the same number
of samples.
% Thus, it is difficult to explain the method's propensity toward the model with
% the smallest parameter space via insufficient prior sampling.
While analyses that sample each model proportional to their parameter space
could be explored, it seems much more likely that the marginal likelihoods
under broad uniform priors prefer simpler models.

As discussed by \citet{Oaks2012}, a straightforward prediction of marginal
likelihoods causing the preference for simple models is that the bias should
disappear as the model generating the data converges to the prior.
\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
assess the model-choice behavior of \msb when the prior model is correct.
They find that the bias actually switches direction (at least for the
regression-adjusted estimates) and the method tends to underestimate the
probability of the model with one divergence event (see Figure 4 of
\citet{Oaks2012}).
The same prediction is not as straightforward for the insufficient-sampling
hypothesis.
Even when the prior is correct, due to the discrete uniform prior on
\numt{} implemented in \msb, models with larger numbers of divergence
events (and thus greater parameter space) will still be less densely
sampled than those with fewer divergence events \citep{Oaks2012}.
Thus, the results of the simulations of \citet{Oaks2012} are more consistent
with larger marginal likelihoods of models with fewer divergence-time
parameters.

Taken together, the results presented here and in \citet{Oaks2012} support
greater marginal likelihoods of simple models under broad priors as a primary
mechanism by which \msb prefers models with clustered divergence times.
Nonetheless, sampling error will always be present in any numerical
Bayesian approximation method.

@@

%CWL: This seems to nit-picky. Probably not worth including it

>>>>

%CWL: I would suggest removing this section
\section*{The use of power analysis to guide applications of \msb}
\citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
divergence-time prior of 0--1 coalescent units ago.
They found that under these prior conditions \msb has the power, assuming a
per-site rate of $1.92\e{-8}$ mutations per generation, to detect multiple
divergence events among 18 taxa when the true divergences were random over
hundreds of thousands of generations or more.
\citet{Oaks2012} performed similar power analyses under three uniform
divergence-time priors as narrow as 0--5 coalescent units, and found the
method was able to detect multiple events among 22 when divergences were random
over millions of generations.
As recommended by \citet{Oaks2012}, it is important that investigators perform
power analyses to determine the method's power for their dataset, and decide if
\msb has sufficient power to address their hypotheses; in the case of the
Philippines dataset, it did not.
When doing so, it is important to consider what prior conditions are
relevant to their empirical system.
A divergence-time prior of 0--5 coalescent units is quite narrow considering
that this expresses the prior belief that all 22 taxon pairs diverged within
this window.
Certainly, it is a rare for there to be enough \emph{a priori} information to
be certain that all taxa diverged within the last 4$\globalpopsize$ generations
(i.e., 0--1 coalescent units).
Also, it seems unlikely that when such prior information is available that
being able to detect multiple divergences on the scale of hundreds of thousands
of generations will add much insight about the evolutionary history of the
taxa.
Assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), with such
prior information the method can reliably reject one divergence event when
divergences are random over a period of 3 million generations.
However, as noted by \citet{Oaks2012}, inferring more than one divergence time
shared across all taxa does not confirm the method is working well when
analyzing data generated under random temporal variation in divergences (e.g.,
an inference of two divergence events could be biogeographically interesting
yet spurious).
Investigators should not limit their assessment of method's power to only
differentiating inferences of one event or more (i.e. $\numt{} = 1$ versus
$\numt{} > 1$).
Rather, looking at the distribution of estimates, as in \citet{Oaks2012},
provides much more information about the behavior of the method.

\section*{Model likelihoods or insufficient prior sampling as the cause of bias
in \msb}
% One of the points of disagreement between \citet{Hickerson2013} and
% \citet{Oaks2012} is the underlying mechanism causing the bias toward models
% with clustered divergences (i.e., models with few divergence events).
It is important to determine the mechanism by which broad uniform priors cause
\msb to support clustered models of divergence.
It is well established that vague priors can be problematic in Bayesian model
selection.
Models that integrate over more parameter space characterized by low
probability of producing the data and relatively high prior density will have
smaller marginal likelihoods \citep{Jeffries1935,Lindley1957}.
Given the uniformly distributed priors on divergence times (and other nuisance
parameters) employed in \msb, models with more divergence parameters will be
forced to integrate over \emph{much} greater parameter space, all with equal
prior density, and much of it with low likelihood.
Thus, it is not surprising in light of this fundamental statistical issue that
the method tends to support simple models.

However, \citet{Hickerson2013} conclude that the bias is due to insufficient
prior sampling.
They argue the widest of the three priors used by \citet{Oaks2012} would
infrequently produce samples with many independent young divergence times.
However, there are some issues with their argument.
The probabilities they present assume the gene divergence times
estimated by \citet{Oaks2012} are correct.
The sole purpose of the estimates of ultrametric gene trees presented by
\citet{Oaks2012} was to provide a very rough comparison of the gene divergence
times across the 22 taxa.
These analyses assumed an arbitrary strict mutation rate of $2\e{-8}$ for all
taxa, and are, of course, subject to estimation error.
Furthermore, the branch-length units of the gene trees are in millions of
years, whereas the divergence time prior of \msb is in generations, thus
\citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
have a generation time of one year.
The argument of \citet{Hickerson2013} that divergence time estimates of
\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
questionable, especially given our findings presented above regarding the
behavior of \msb when empirically informed uniform priors are employed.

If we make all of these assumptions and assume the gene divergence times are
estimated without error, the Hickerson et al.'s \citeyear{Hickerson2013}
argument only applies to one of the three different priors used by
\citet{Oaks2012}.
The narrowest prior on divergence times used by \citet{Oaks2012} ($U(0, 5
mya)$) closely mirrors the range of estimates of gene-divergence times.
Applying Hickerson et al.'s (\citeyear{Hickerson2013}) probability equations
demonstrates this prior is densely populated with samples with large numbers of
divergence parameters with values younger than the estimated gene divergence
estimates.
Thus, if insufficient prior sampling is to blame for the bias, it should be
much reduced under the narrow prior on \divt{}.
However, the magnitude of the bias is very similar across all three priors
explored by \citet{Oaks2012}.
\citet{Hickerson2013} point out a case where the narrow prior performs
slightly better (panel L of Figures S32, S37, and S38 of \citet{Oaks2012}).
However, it is important to note that these results suffered from a bug
in \msb, and there are many cases after \citet{Oaks2012} corrected the 
bug where the narrow prior performs slightly worse (see panels D--J of
Figures 3 and S12).

To disentangle whether model likelihoods or insufficient prior sampling is to
blame for the method's spurious support for simple models, we must look at the
different predictions made by these two phenomena.
One example, as discussed by \citet{Oaks2012}, is that insufficient prior
sampling should create higher variance in posterior estimates, and thus it
should cause analyses to be highly sensitive to the number of samples drawn
from the prior.
\citet{Oaks2012} do not see such sensitivity when they compare prior sample
sizes of $2\e6$, $5\e6$, and $10^7$.

To explore this prediction further, we repeat the analysis of the Philippines
dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim U(0,
10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{} \sim
(0.0005, 0.02)$), using a very large prior sample size of $10^8$.
When we look at the trace of the estimates of the dispersion index of
divergence times (\vmratio{}) as the prior samples accumulate
(Figure~\ref{figSamplingError}) we see no trend in either the unadjusted or
GLM-regression-adjusted estimates.
While sampling error is always a reality, clearly it is not playing
a large role in the results of \citet{Oaks2012} or presented above.

As discussed by \citet{Oaks2012}, a straightforward prediction of marginal
likelihoods causing the preference for simple models is that the bias should
disappear as the model generating the data converges to the prior.
\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
assess the model-choice behavior of \msb when the prior model is correct.
The results confirm the prediction as \msb actually tends to underestimate
the probability of the one-divergence model (see Figure 4 of
\citet{Oaks2012}).
These results are not clearly predicted if insufficient-sampling was causing
the bias.
Even when the prior is correct, due to the discrete uniform prior on
\numt{} implemented in \msb, models with larger numbers of divergence
events (and thus greater parameter space) will still be less densely
sampled than those with fewer divergence events \citep{Oaks2012}.
Thus, the results of the simulations of \citet{Oaks2012} are more consistent
with larger marginal likelihoods of models with fewer divergence-time
parameters.

Furthermore, the results above demonstrate the model-averaging
approach of \citet{Hickerson2013} is prefers models with narrower \divt{}
priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
\ref{figPowerOmegaProb}).
% This further suggests the primary cause is the uniform
% priors reducing the marginal likelihoods of large models.
In all of these analyses, each of the prior models have the same number of
samples.
While analyses that sample each model proportional to their parameter space
could be explored, it clear that the marginal likelihoods under broad uniform
priors will be greater for simpler models.
% Thus, it is difficult to explain the method's propensity toward the model with
% the smallest parameter space via insufficient prior sampling.
Taken together, the results presented here and in \citet{Oaks2012} support
greater marginal likelihoods of simple models under broad priors as a primary
mechanism by which \msb prefers models with clustered divergence times.
Nonetheless, sampling error will always be present in any numerical
Bayesian approximation method.

@@
